---
title: "Simulation"
author: "Ying Jin"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = '')

set.seed(625)

library(tidyverse)
library(here)
library(kableExtra)

load(here("Data/SimOutput/SimOutput_J200.RData"))
load(here("Data/SimOutput/SimOutput_J200_ref.RData"))


# valid_id <- which(pred_time_vec<=10)

```

# Simulation set-up

This file documents the simulation scheme of multi-level functional data. 

- Sample size: training = 100, test = 100
- Number of visits J = 3
- Number of observations at each vitis K = 200 (when K = 100, numeric issues are likely to happen)
- Make prediction with observations until the middle of visit 2 and 3 separately
- 100 simulations



- Binary outcomes: 

$$Y_{ij}(t) \sim Bernoulli(g(\eta_{ij}(t)))$$

$$\eta_{ij}(t)=b_0(t)+\sum_{l=1}^4\xi_{il}\phi_l(t)+\sum_{m=1}^4\zeta_{ijm}\psi_m(t)$$


- Level 1: 

$$\phi_l(t) = \{\sqrt{2}sin(2\pi t), \sqrt2 cos(2\pi t), \sqrt2 sin (4\pi t), \sqrt2 cos(4\pi t)\}$$

$$\xi_{il} \sim N(0, \lambda_l), \hspace{0.5cm} \lambda_l = 0.5^{l-1}, l=1,2, 3, 4$$

# Results

## Figure

```{r}
pred_list_allM[[1]] %>%
  filter(id %in% 101:104) %>%
  mutate(eta_pred_J1 = ifelse(t<=0.5, NA, eta_pred_J1),
         eta_pred_J2 = ifelse(visit==1|(visit %in% 2:3 & t<=0.5), NA, eta_pred_J2),
         eta_pred_J3 = ifelse(visit %in% 1:2|(visit==3&t<=0.5), NA, eta_pred_J3)) %>%
  mutate_at(vars(starts_with("eta_pred")), function(x)(exp(x)/(1+exp(x)))) %>% 
  ggplot()+
  geom_line(aes(x=t, y=probs, col="True"))+
  geom_line(aes(x=t, y=eta_pred_J1, col = "No full visits"), linetype="dashed", 
            na.rm = T)+
  geom_line(aes(x=t, y=eta_pred_J2, col = "1 full visit"), linetype="dashed", 
            na.rm = T)+
  geom_line(aes(x=t, y=eta_pred_J3, col = "2 full visits"), linetype="dashed", 
            na.rm = T)+
  geom_point(aes(x=t, y=Y), size = 0.5)+
  facet_grid(row = vars(id), col=vars(visit))+
  labs(title = "gmFPCA")
```


```{r}
pred_list_allM_ref[[1]] %>%
  filter(id %in% 101:104) %>%
  mutate(eta_pred_J1 = ifelse(t<=0.5, NA, eta_pred_J1),
         eta_pred_J2 = ifelse(visit==1|(visit %in% 2:3 & t<=0.5), NA, eta_pred_J2),
         eta_pred_J3 = ifelse(visit %in% 1:2|(visit==3&t<=0.5), NA, eta_pred_J3)) %>%
  mutate_at(vars(starts_with("eta")), function(x)(exp(x)/(1+exp(x)))) %>% 
  ggplot()+
  geom_line(aes(x=t, y=eta, col="True"))+
  geom_line(aes(x=t, y=eta_pred_J1, col = "No full visits"), linetype="dashed", 
            na.rm = T, alpha = 0.5, linewidth = 1)+
  geom_line(aes(x=t, y=eta_pred_J2, col = "1 full visit"), linetype="dashed",
            na.rm = T, alpha = 0.5, linewidth = 0.8)+
  geom_line(aes(x=t, y=eta_pred_J3, col = "2 full visit"), linetype="dashed",
            na.rm = T, alpha = 0.5, linewidth = 0.6)+
  geom_point(aes(x=t, y=Y), size = 0.5)+
  facet_grid(row = vars(id), col=vars(visit))+
  labs(title = "GLMMadaptive")
```

# Computation time

```{r}
# print("Computation time:")
data.frame("gmFPCA" = c(mean(fit_time_vec), 
                        mean(pred_time_vec)),
           "GLMMadaptive" = c(mean(fit_time_vec_ref),
                              mean(pred_time_vec_ref))) %>% 
  mutate("Time (mins)" = c("Fit", "Prediction"), .before=1) %>%
  kable(digits = 2) %>% 
  kable_styling(full_width = F)



```


# Integerated squared error

- Over the entire prediction window

```{r, message=FALSE}
ise_gmfpca <- bind_rows(pred_list_allM, .id = "iter") %>% 
  mutate(window = cut(t, breaks = c(0, 0.5, 1), include.lowest = T)) %>%
  mutate(error_J1 = (eta_pred_J1-eta)^2,
         error_J2 = (eta_pred_J2-eta)^2,
         error_J3 = (eta_pred_J3-eta)^2) %>%
  group_by(iter, visit, window, id) %>%
  summarise_at(vars(starts_with("error_")), sum) %>% 
  group_by(iter, visit, window) %>%
  summarise_at(vars(starts_with("error_")), mean) %>% 
  group_by(visit, window) %>%
  summarise_at(vars(starts_with("error_")), mean)

ise_gmfpca <- ise_gmfpca %>% filter(window == "(0.5,1]")
ise_gmfpca$error_J2[1] <- NA
ise_gmfpca$error_J3[1:2] <- NA
colnames(ise_gmfpca) <- c("Visit", "Time",
                          "All half visits", "Full visit 1, half visit 2 and 3",
                          "Full visit 1 and 2, half visit 3")
```


```{r}
ise_glmmad <- bind_rows(pred_list_allM_ref, .id = "iter") %>% 
  mutate(window = cut(t, breaks = c(0, 0.5, 1), include.lowest = T)) %>%
  mutate(error_J1 = (eta_pred_J1-eta)^2,
         error_J2 = (eta_pred_J2-eta)^2,
         error_J3 = (eta_pred_J3-eta)^2) %>%
  group_by(iter, visit, window, id) %>%
  summarise_at(vars(starts_with("error_")), sum) %>% 
  group_by(iter, visit, window) %>%
  summarise_at(vars(starts_with("error_")), mean) %>% 
  group_by(visit, window) %>%
  summarise_at(vars(starts_with("error_")), mean)  

ise_glmmad <- ise_glmmad %>% filter(window == "(0.5,1]")
ise_glmmad$error_J2[1] <- NA
ise_glmmad$error_J3[1:2] <- NA
colnames(ise_glmmad) <- c("Visit", "Time",
                          "All half visits", "Full visit 1, half visit 2 and 3",
                          "Full visit 1 and 2, half visit 3")
```

```{r}
# print("Integrated squared error:")
ise_gmfpca %>% 
  add_column(ise_glmmad[,3:5], .name_repair = "minimal") %>%
  kable(digits = 2) %>% 
  kable_styling(full_width = F) %>%
  add_header_above(c("Prediction window" = 2,
                         "Observed window" = 6))
```

For visit 3, the predictions are identical. This is not a coding error. Because prediction of the three visits are made at the same time, and the other visits are not identical. 
